---
title: "R in Action - Chapter 4:Basic data Mangement"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##### Create leadership data frame
```{r}
manager <- c(1,2,3,4,5)
date <- c("10/24/08","10/28/08","10/1/08","10/12/08","5/1/09")
country <- c("US","US","UK","UK","UK")
gender <- c("M","F","F","M","F")
age <- c(32,45,25,39,99)
q1 <- c(5,3,3,3,2)
q2 <- c(4,5,5,3,2)
q3 <- c(5,2,5,4,1)
q4 <- c(5,5,5,NA,2)
q5 <- c(5,5,2,NA,1)
leadership <- data.frame(manager,date,country,gender,age,q1,q2,q3,q4,q5,
stringsAsFactors = FALSE)
```

##### Recoding variables
```{r}
leadership$age[leadership$age == 99] <- NA
leadership$agecat[leadership$age > 75] <- "Elder"
leadership$agecat[leadership$age  <= 75 & leadership$age >= 55] <- "Middler Age"
leadership$agecat[leadership$age  <55] <- "Young"
```
Same as use `within()` function:  
```{r}
leadership <- within(leadership,{
age[age == 99] <- NA
agecat <- NA
agecat[age > 75] <- "Elder"
agecat[age >= 55 & age <= 75] <- "Middler Age"
agecat[age < 55] <- "Young"
}
)
library(plyr)
leadership <- rename(leadership,c(manager = "managerID",date = "testDate",
q1 = "item1",q2 = "item2",q3 = "item3",q4 = "item4",q5 = "item5"))
```
##### Missing values
Use `na.omit()` deletes any rows with missing data (only with small amount of data is missing)
```{r}
newdata <- na.omit(leadership)
```
##### Date Variable
| Symbol | Meaning               | Example |
|--------|-----------------------|---------|
| %d     | Day as number         | 01-31   |
| %a     | Abbreviated weekday   | Mon     |
| %A     | Unabbreviated weekday | Monday  |
| %m     | Month                 | 00-12   |
| %b     | Abbreviated month     | Jan     |
| %B     | Unabbreviated month   | Janary  |
| %y     | Two-digit year        | 07      |
| %Y     | Four-digit year       | 2007    |

```{r}
leadership$testDate <- as.character(as.Date(leadership$testDate,"%m/%d/%y"))
```
##### Sort dataset 
Use `order()` to sort, by default is ascending
```{r}
newdata <- leadership[order(leadership$age),]
newdata
```
##### combine dataset
1.`merge`: merge data frames horizontally,`by` common keys;
Or can use `cbind` without specify common keys.   
2.`rbind`:join data frames vertically,add rows to data frames.
##### Use `subset()` function 
```{r}
subset(leadership,age >= 35|age < 24,select = c(item1,item2,item3,item4))
```
##### Random Sample
Use `sample()` function: sample(x,size,replace = FALSE,prob = NULL)
```{r}
leadership[sample(1:nrow(leadership),3,replace = FALSE),]
```
##### use SQL
Use `sqldf` package
```{r}
library(sqldf)
newdf <- sqldf(
  "select * from mtcars where carb = 1 order by mpg",
  row.names = TRUE
)
newdf
sqldf("select avg(mpg) as avg_mpg,avg(disp) as avg_disp,gear from mtcars where cyl in (4,6) group by gear")
```


